\Title

### 2021-10-12 latex/paper/main.tex
Final edits of the paper to send to Manfredi and Xingbo

\usepackage{generic}
%/========== Preamble ==========/%
% Packages required by ieeecolor
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{xspace}
\usepackage{enumerate}
%---------- More packages ----------%
% NOTE: cannot use amsthm for some reason
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}

@@

%/========== Abstract ==========/%
\begin{abstract}
    In this article we study virtual nonholonomic constraints, 
    relations between the generalized coordinates and momenta of a
    mechanical system that can be enforced via feedback control.
    We design a constraint which emulates gymnastics giant motion in an
    acrobot, and prove that this constraint can inject or dissipate
    energy based on the sign of a design parameter.
    The proposed constraint is tested both in simulation and experimentally on a real-world acrobot,
    demonstrating highly effective energy regulation properties and robustness
    to a variety of disturbances.
\end{abstract}

@@

Previous attempts at acrobot giant generation have involved
trajectory tracking, partial feedback linearization, or other energy-based
methods
( see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While all these approaches succeed at making the acrobot rotate around the
bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
That is, none of these leg controllers track a function of the acrobot's body
angle and velocity.
In 2016, Wang designed a controller \cite{xingbo_thesis} which tracked the
body angle and an \textit{estimate} of the velocity, but not the velocity itself.
His approach was a preliminary version of a recent technique know as the method
of virtual nonholonomic constraints \cite{hybrid_zero_dynamics_bipedal_nhvcs}.

@@

In this article, we design a virtual nonholonomic constraint which provably
injects energy into the acrobot through human-like giant motion.
In particular, any acrobot constrained by this \vnhc is guaranteed to perform
rotations around the bar.
Under suitable conditions on the acrobot's physical parameters, this \vnhc
also enables the acrobot to rotate around the bar at a desired speed.
We will provide simulations to show how one can use \vnhcs to regulate energy,
along with experimental results which demonstrate the robustness of this
behaviour to model uncertainty, sensor noise, and a variety of external
disturbances.

@@

Our goal is to design a smooth function \(f : \Sone \times \R\) such that the relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian} can be enforced asymptotically via feedback control (in Section \ref{sec:vnhc} we call this a  virtual nonholonomic constraint).   We will further require that the dynamics of the acrobot, when the relation holds, gain or lose energy in a sense that will be defined precisely in Section~\ref{sec:energy-inject-diss}.

@@

Since the control input \(\tau\) only affects the actuated momentum,
the system above is already in simply actuated form.
Its state space is \(\Sone \times \Re \times \Re \times \Re\).
We can therefore apply the theory from Section \ref{sec:vnhc} to design
a \vnhc of the form \(q_a = f(q_u,p_u)\) (i.e., a \vnhc 
\(h(q,p_u) = q_a - f(q_u,p_u) = 0\)).
Since we need the \vnhc to be regular, the following proposition will be
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    Any relation \(q_a = f(p_u)\) 
    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
    \vnhc of order 1 for the acrobot in \eqref{eqn:acrobot-hamiltonian}.
\end{prop}
\begin{proof}
    The decoupling matrix \eqref{eqn:decoupling-matrix} for the acrobot
    evaluates to
    \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
    Since \(\partial_{q_u} f = 0\), this function is positive for all values
    of \(q_a\), and hence is full rank \(1\) everywhere on the constraint manifold.
\end{proof}

@@

\subsection{Energy Injection and Dissipation with the Proposed \vnhc}\label{sec:energy-inject-diss}
Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
that the legs stay fully extended.
The acrobot becomes a nominal pendulum with two masses, whose total mechanical
energy is
\begin{equation}\label{eqn:nominal-energy}
    E(q_u,p_u) \eqdef \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    .
\end{equation}
The function $E(q_u,p_u)$ is a first integral for the constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics}, i.e., each solution of the constrained dynamics is confined to a level set of $E$, and each orbit not containing equilibria is a connected component of a level set of $E$. Given scalars $R$ and $R_1<R_2$, and letting $\# \in \{$ ``$<$''$,$''$\leq$''$,$''$>$''$,$''$\geq$''$\}$, we define the following sets (see Figure~\ref{fig:domains}):
\begin{equation}\label{eq:E_sets}
\begin{aligned}
E_R &\eqdef \{(q_u,p_u): E(q_u,p_u) = R\}, \\
E_{\# R} &\eqdef \{(q_u,p_u): E(q_u,p_u) \#R\}, \\
E_{[R_1,R_2]} &\eqdef \{(q_u,p_u): R_1 \leq E(q_u,p_u) \leq R_2\}.
\end{aligned}
\end{equation}
%
%
%
\begin{figure}[htb]
    \centering
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_EleqR.pdf}
        \caption{The set $E_{\leq R}$ with $0 < R < \bar R$ and an oscillation gaining energy.} 
        \label{fig:acrobot-oscillation-domain}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_ER1R2.pdf}
        \caption{The set $E_{[R_1,R_2]}$, with $\bar R < R_1 < R_2$, and a rotation gaining energy.}
            \label{fig:acrobot-o2}
    \end{subfigure}
    \caption{Oscillations and rotations gaining energy. 
}
\label{fig:domains}
\end{figure}
%
%
%

@@

\begin{rem}
Theorem~\ref{thm:acrobot-oscillations} concerning oscillations gai\-ning/lo\-sing energy states that for any set of physical parameters, the acrobot constrained by
\eqref{eqn:acrobot-constraint} will gain enough energy to exit in finite time a level set of the function $E$ arbitrarily close to the boundary $E_{\bar R}$ separating oscillations from rotations, provided the parameter $I$ in~\eqref{eqn:acrobot-constraint} is chosen small enough. The result is semiglobal relative to the set $E_{<\bar R}$ in that energy gain can be achieved on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$ by a suitably small $I$.
%% Moreover, after this finite time, as long as the solution is contained in the set $E_{< \bar R}$, then its intersections with the Poincar\'e section $\cP_o$ will occur outside of $E_{<R}$. Roughly speaking, this means the oscillations will retain the energy they have acquired.
Vice versa, for small {\em negative} $I$, the oscillations will lose energy on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$.

@@

We will now simulate the effects of constraining the physical acrobot with the \vnhc \eqref{eqn:acrobot-constraint}, thereby demonstrating that \vnhcs are robust to model mismatch. According to Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, the control parameter \(I\) must be ``small" for the \vnhc~\eqref{eqn:acrobot-constraint} to inject (or dissipate) energy into the simplified acrobot. The Theorem does not specify how small \(|I|\) must be; while we could make it arbitrarily small in simulations, we will eventually implement this \vnhc on a physical testbed where \(|I|\) must be large enough to overcome friction.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
    \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/6703e828acdcd38058ca884d13660a88}}
    \label{fig:acrobot-unperturbed-orbit}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping. A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/164a21f0a52d4d56ef60bf1911545e0e}}
    \label{fig:acrobot-stopped-orbit}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forward push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing. A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/b13d4dc9f6daefa7f4d56dbf1b1bd482.}}
\end{figure}

@@

By properties~\eqref{eq:lemma_assumptions:1} and~\eqref{eq:lemma:assumptions:3} and since $f_r, f_\theta$ are continuous and $K$ is compact, there exist $I_1,\varepsilon_1,\varepsilon_2>0$ such that for each $(r,\theta) \in K$ and $I \in [-I_1,I_1]$,

@@

&f_r(r,\theta,I) > \varepsilon_1,\label{eq:properties:new_coords:1}\\

@@

Property~\eqref{eq:properties:new_coords:1} implies that for $I \in [-I_1,I_1]$ the first component of any solution $(r(t),\theta(t))$ contained in $K$ is a strictly monotonic function of $t$, and therefore $r(t)$  can be expressed as $r(t) = \hat r(\theta(t))$, where $\hat r(\theta)$ is a solution of

@@

    + \Delta(\theta,r_0,I) \\

@@

where the function $\Delta(r,\theta,I)$ is continuous and $I \mapsto \Delta(r,\theta,I)$ is $O(I^2)$.

@@

Now consider the Poincar\'e section $\cP = \{(r,\theta) \in \cV: \theta =0\}$ defined in the lemma statement. Using~\eqref{eq:Taylor_exp}, and~\eqref{eq:ronehat_solution}, the Poincar\'e map on $\cP$ is $P: \cP \to \Re_{>0}$, 

@@

Define $\cU \eqdef E_{<\bar R} \setminus \{0\}$ and $\cV \eqdef \mathopen]0,\pi\mathclose[$, and consider the coordinate transformation\footnote{Despite the presence of a $\text{sgn}(\cdot)$ function, the coordinate transformation $T_1$ is actually smooth.} $T: \cU \to \cV$ defined as
\begin{equation}\label{eq:T1}
\begin{aligned}
T: & \, (q_u,p_u) \mapsto (r,\theta)  \\
%
r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right) \\
%
\theta &=\left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = \rone(q_u,p_u)},
\end{aligned}
\end{equation}
whose inverse is
\[
\begin{aligned}
T^{-1}: & \, (r,\theta) \mapsto (q_u,p_u)\\
%
q_u &= r \cos(\theta) \\
% 
p_u &=-\sign{\sin(\theta)} 
        \sqrt{30m^2gl^3\cos(r\cos(\theta) - \cos(r))}.
\end{aligned}
\]
It is easily seen that for each $R \in \mathopen ]0,\bar R\mathclose[$, $T(E_R) = \{(r,\theta)\in \cV : r = \acos(1 - R/(3 m g l))\}$, and the function $R \mapsto \acos(1 - R/(3 m g l))$ is monotonically increasing. The diffeomorphism $T: \cU \to \cV$ therefore satisfies the first requirement of Lemma~\ref{lem:gain_lose} for any choice of $0< R_1 < R_2 < \bar R$.

In $(r,\theta)$ coordinates, $T$ maps system~\eqref{eqn:acrobot-constrained-dynamics} into a system of the form~\eqref{eq:r_theta_system}, where is straightforward to show that \(f_r(\theta,r,0) \equiv 0\)
and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}.
\end{equation}
In the above we used the notation \(c_\theta \eqdef \cos(\theta)\), \(s_\theta \eqdef \sin(\theta)\), \(c_r \eqdef \cos(r)\), and \(s_r \eqdef \sin(r)\). We note that~\eqref{eqn:theta-dot-nom-osc} has removable singularities at \(\theta \in \{0,\pi\}\), since taking the limits as \(\theta\to 0,\pi\) gives
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
and these limits smooth and well-defined for all \(r \in \, ]0,\pi[\). From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit}, one can verify that \(f_\theta(\theta,r,0) > 0\) for all $(\theta,r) \in \cV$, and thus assumption~\eqref{eq:lemma_assumptions:1} holds. Letting \(g= f_r/f_\theta\), one can also verify that \((\partial g/\partial \hat r)(\theta,\hat r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. Symbolic computations reveal that \(\partial_I g(r,\theta,I) \big|_{I=0} = L a(r,\theta)\), where
\begin{align*}
    L &\eqdef \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(r,\theta) &\eqdef \frac{
        r |s_\theta| \left(
        5 c_{r} \cos(r c_\theta) - 8 \cos(r c_\theta)^2 + 3
    \right)
    }{
    s_{r}\sqrt{\cos(r c_\theta) - c_{r}}
    }
    .
\end{align*}
%
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\linewidth]{Qro.png}
    \caption{The graph of \(r \mapsto \int_0^{2\pi} a(r,\theta) d\theta\).}
    \label{fig:acrobot-Q}
\end{figure}
%
Notice that \(L\) is a positive constant which depends only on \(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(r,\theta)\) is adimensional. We have
\[
\int_0^{2\pi} \partial_I g(r,\theta,I) \big|_{I=0} d \theta =L \int_0^{2\pi} a(r,\theta) d\theta.
\]
The graph of \(\int_0^{2\pi} a(r,\theta) d\theta\), shown in
Figure \ref{fig:acrobot-Q}, indicates that for each $r \in \mathopen] 0,\pi\mathclose[$, the function $r \mapsto \int_0^{2\pi} a(r,\theta) d\theta$ is positive, and assumption~\eqref{eq:lemma:assumptions:3} holds. Thus the hypotheses of Lemma~\ref{lem:gain_lose} are satisfied, proving the theorem.
\hfill\QED

@@

In $(r,\theta)$ coordinates, both $T^+$ and $T^-$ map system~\eqref{eqn:acrobot-constrained-dynamics} to a system of the form~\eqref{eq:r_theta_system}, where one can verify that
\(f_r(\theta,r,0) \equiv 0\) and
\[
 f_\theta(\theta,r,0) = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
,
\]
which is positive so assumption~\eqref{eq:lemma_assumptions:1} holds. Letting $g = f_r / f _\theta$, one can verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. The function \(b(r,\theta) = (\partial g/\partial I)(r,\theta,0)\) is given in the statement of Theorem~\ref{thm:acrobot-rotations}, and by assumption we have that condition~\eqref{eq:lemma:assumptions:3} of Lemma~\ref{lem:gain_lose} holds. Theorem~\ref{thm:acrobot-rotations} now follows from  Lemma~\ref{lem:gain_lose} and the fact that $\cU^+ \cup \cU^- = E_{>\bar R}$.
\hfill \QED

>>>>

\usepackage{generic}
\usepackage{hyperref}
%/========== Preamble ==========/%
% Packages required by ieeecolor
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{xspace}
\usepackage{enumerate}
%---------- More packages ----------%
% NOTE: cannot use amsthm for some reason
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}

@@

%/========== Abstract ==========/%
\begin{abstract}
    In this article we study virtual nonholonomic constraints, which are
    relations between the generalized coordinates and momenta of a
    mechanical system that can be enforced via feedback control.
    We design a constraint which emulates gymnastics giant motion in an
    acrobot, and prove that this constraint can inject or dissipate
    energy based on the sign of a design parameter.
    The proposed constraint is tested both in simulation and experimentally on a real-world acrobot,
    demonstrating highly effective energy regulation properties and robustness
    to a variety of disturbances.
\end{abstract}

@@

Previous attempts at acrobot giant generation have involved
trajectory tracking, partial feedback linearization, or other energy-based
methods
( see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While all these approaches succeed at making the acrobot rotate around the
bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
That is, none of these leg controllers track a function of the acrobot's body
angle and velocity.
In 2016, Wang designed a controller which tracked the
body angle and an \textit{estimate} of the velocity, but not the velocity itself
\cite{xingbo_thesis}.
His approach was a preliminary version of a recent technique know as the method
of virtual nonholonomic constraints.

@@

In this article, we design a virtual nonholonomic constraint which provably
injects energy into the acrobot through human-like giant motion.
%In particular, any acrobot constrained by this \vnhc is guaranteed to perform
%rotations around the bar.
%Under suitable conditions on the acrobot's physical parameters, this \vnhc
%also enables the acrobot to rotate around the bar at a desired speed.
We will provide simulations to show how one can use \vnhcs to regulate energy,
along with experimental results which demonstrate the robustness of this
behaviour to model uncertainty, sensor noise, and a variety of external
disturbances.

@@

Our goal is to design a smooth function \(f : \Sone \times \R \to \Sone\) such that the
relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian} can be
enforced asymptotically via feedback control (in Section \ref{sec:vnhc} we call
this a  virtual nonholonomic constraint).   We will further require that the
dynamics of the acrobot, when the relation holds, gain or lose energy in a sense
that will be defined precisely in Section~\ref{sec:energy-inject-diss}.

@@

Since the control input \(\tau\) only affects the actuated momentum,
the system above is already in simply actuated form.
Its state space is \(\Sone \times \Sone \times \Re \times \Re\).
We can therefore apply the theory from Section \ref{sec:vnhc} to design
a \vnhc of the form \(q_a = f(q_u,p_u)\) (i.e., a \vnhc 
\(h(q,p_u) = q_a - f(q_u,p_u) = 0\)).
Since we need the \vnhc to be regular, the following proposition will be
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    Any relation \(q_a = f(p_u)\) 
    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
    \vnhc of order 1 for the acrobot in \eqref{eqn:acrobot-hamiltonian}.
\end{prop}
\begin{proof}
    The decoupling matrix \eqref{eqn:decoupling-matrix} for the acrobot
    evaluates to
    \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
    Since \(\partial_{q_u} f = 0\), this function is positive for all values
    of \(q_a\), and hence is full rank \(1\) everywhere on the constraint manifold.
\end{proof}

@@

\subsection{Energy Injection and Dissipation with the Proposed \vnhc}\label{sec:energy-inject-diss}
Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
that the legs stay fully extended.
The acrobot becomes a nominal pendulum with two masses, whose total mechanical
energy is
\begin{equation}\label{eqn:nominal-energy}
    E(q_u,p_u) \eqdef \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    .
\end{equation}
The function $E(q_u,p_u)$ is a first integral for the constrained
dynamics~\eqref{eqn:acrobot-constrained-dynamics}, i.e., each solution of the
constrained dynamics is confined to a level set of $E$, and each orbit not
containing equilibria is a connected component of a level set of $E$. Given
scalars $R$ and $R_1<R_2$, and letting $\# \in \{$
``$<$''$,$``$\leq$''$,$``$>$''$,$``$\geq$''$\}$, we define the following sets
(see Figure~\ref{fig:domains}):
\begin{equation}\label{eq:E_sets}
\begin{aligned}
E_R &\eqdef \{(q_u,p_u): E(q_u,p_u) = R\}, \\
E_{\# R} &\eqdef \{(q_u,p_u): E(q_u,p_u) \#R\}, \\
E_{[R_1,R_2]} &\eqdef \{(q_u,p_u): R_1 \leq E(q_u,p_u) \leq R_2\}.
\end{aligned}
\end{equation}
%
%
%
\begin{figure}[htb]
    \centering
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_EleqR.pdf}
        \caption{The set $E_{\leq R}$ with $0 < R < \bar R$ and an oscillation gaining energy.} 
        \label{fig:acrobot-oscillation-domain}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\linewidth}
        \includegraphics[width=\linewidth]{acrobot_ER1R2.pdf}
        \caption{The set $E_{[R_1,R_2]}$, with $\bar R < R_1 < R_2$, and a rotation gaining energy.}
            \label{fig:acrobot-o2}
    \end{subfigure}
    \caption{Oscillations and rotations gaining energy. 
}
\label{fig:domains}
\end{figure}
%
%
%

@@

\begin{rem}
Theorem~\ref{thm:acrobot-oscillations} concerning oscillations
gai\-ning/lo\-sing energy states that for any set of physical parameters, the
acrobot constrained by
\eqref{eqn:acrobot-constraint} will gain enough energy to exit (in finite time) a
level set of the function $E$ arbitrarily close to the boundary $E_{\bar R}$
separating oscillations from rotations, provided the parameter $I$
in~\eqref{eqn:acrobot-constraint} is chosen small enough. The result is
semiglobal relative to the set $E_{<\bar R}$ in that energy gain can be achieved
on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$ by a suitably small $I$.
%% Moreover, after this finite time, as long as the solution is contained in the set $E_{< \bar R}$, then its intersections with the Poincar\'e section $\cP_o$ will occur outside of $E_{<R}$. Roughly speaking, this means the oscillations will retain the energy they have acquired.
Vice versa, for small {\em negative} $I$, the oscillations will lose energy on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$.

@@

We will now simulate the effects of constraining the physical acrobot with the
\vnhc \eqref{eqn:acrobot-constraint}, thereby demonstrating that \vnhcs are
robust to model mismatch. According to Theorems~\ref{thm:acrobot-oscillations}
and~\ref{thm:acrobot-rotations}, the control parameter \(I\) must be ``small"
for the \vnhc~\eqref{eqn:acrobot-constraint} to inject (or dissipate) energy
into the simplified acrobot. The theorems do not specify how small \(|I|\) must
be; while we could make it arbitrarily small in simulations, we will eventually
implement this \vnhc on a physical testbed where \(|I|\) must be large enough to
overcome friction.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
    \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at 
    {\footnotesize
\url{https://play.library.utoronto.ca/watch/6703e828acdcd38058ca884d13660a88}}.}
    \label{fig:acrobot-unperturbed-orbit}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red)
    stopping. A video of this experiment is available at {\footnotesize
\url{https://play.library.utoronto.ca/watch/164a21f0a52d4d56ef60bf1911545e0e}}.}
    \label{fig:acrobot-stopped-orbit}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forward push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red)
    pushing. A video of this experiment is available at 
    {\footnotesize\url{https://play.library.utoronto.ca/watch/b13d4dc9f6daefa7f4d56dbf1b1bd482}.}
    }
\end{figure}

@@

By properties~\eqref{eq:lemma_assumptions:1} and~\eqref{eq:lemma:assumptions:3}
and since $f_r, f_\theta$ are continuous and $K$ is compact, there exist
$I_1,\varepsilon_1,\varepsilon_2>0$ such that for each $(r,\theta) \in K$ and $I
\in [-I_1,I_1]$,

@@

&f_\theta(r,\theta,I) > \varepsilon_1,\label{eq:properties:new_coords:1}\\

@@

Property~\eqref{eq:properties:new_coords:1} implies that for $I \in [-I_1,I_1]$
the second component of any solution $(r(t),\theta(t))$ contained in $K$ is a
strictly monotonic function of $t$, and therefore $r(t)$  can be expressed as
$r(t) = \hat r(\theta(t))$, where $\hat r(\theta)$ is a solution of

@@

    + \Delta(r_0,\theta,I) \\

@@

where the function $\Delta(r,\theta,I)$ is continuous and $I \mapsto
\Delta(r,\theta,I)$ is $O(I^2)$.

@@

Now consider the Poincar\'e section $\cP = \{(r,\theta) \in \cV: \theta =0\}$. Using~\eqref{eq:Taylor_exp},
and~\eqref{eq:ronehat_solution}, the Poincar\'e map on $\cP$ is $P: \cP \to
\Re_{>0}$, 

@@

Define $\cU \eqdef E_{<\bar R} \setminus \{0\}$ and $\cV \eqdef
\mathopen]0,\pi\mathclose[ \, \times \Sone$, and consider the coordinate
transformation\footnote{Despite the presence of a $\text{sgn}(\cdot)$ function,
this coordinate transformation $T$ is actually smooth.} $T: \cU \to \cV$
defined as
\begin{equation}\label{eq:T1}
\begin{aligned}
T: & \, (q_u,p_u) \mapsto (r,\theta)  \\
%
r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right) \\
%
\theta &=\left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = \rone(q_u,p_u)},
\end{aligned}
\end{equation}
whose inverse is
\[
\begin{aligned}
T^{-1}: & \, (r,\theta) \mapsto (q_u,p_u)\\
%
q_u &= r \cos(\theta) \\
% 
p_u &=-\sign{\sin(\theta)} 
        \sqrt{30m^2gl^3\cos(r\cos(\theta) - \cos(r))}.
\end{aligned}
\]
It is easily seen that for each $R \in \mathopen ]0,\bar R\mathclose[$, 
$T(E_R) = \{(r,\theta)\in \cV : r = \arccos(1 - R/(3 m g l))\}$, and the function
$R \mapsto \arccos(1 - R/(3 m g l))$ is monotonically increasing. The
diffeomorphism
$T: \cU \to \cV$ therefore satisfies the first requirement of
Lemma~\ref{lem:gain_lose} for any choice of $0< R_1 < R_2 < \bar R$.

In $(r,\theta)$ coordinates, $T$ maps
system~\eqref{eqn:acrobot-constrained-dynamics} into a system of the
form~\eqref{eq:r_theta_system}, where is straightforward to show that
\(f_r(r,\theta,0) \equiv 0\)
and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    f_\theta(r,\theta,0) = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}.
\end{equation}
In the above we used the notation \(c_\theta \eqdef \cos(\theta)\), \(s_\theta \eqdef \sin(\theta)\), \(c_r \eqdef \cos(r)\), and \(s_r \eqdef \sin(r)\). We note that~\eqref{eqn:theta-dot-nom-osc} has removable singularities at \(\theta \in \{0,\pi\}\), since taking the limits as \(\theta\to 0,\pi\) gives
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}f_\theta(r,\theta,0)
    = \lim\limits_{\theta \to \pi} f_\theta(r,\theta,0)
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
and these limits are smooth and well-defined for all \(r \in \, ]0,\pi[\). From
\eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit}, one can verify that
\(f_\theta(r,\theta,0) > 0\) for all $(r,\theta) \in \cV$, and thus
assumption~\eqref{eq:lemma_assumptions:1} holds. Letting \(g= f_r/f_\theta\),
one can also verify that \(\partial_r g(r,\theta,0) \equiv
0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. Symbolic computations
reveal that \(\partial_I g(r,\theta,I) \big|_{I=0} = L a(r,\theta)\), where
\begin{align*}
    L &\eqdef \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(r,\theta) &\eqdef \frac{
        r |s_\theta| \left(
        5 c_{r} \cos(r c_\theta) - 8 \cos(r c_\theta)^2 + 3
    \right)
    }{
    s_{r}\sqrt{\cos(r c_\theta) - c_{r}}
    }
    .
\end{align*}
%
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\linewidth]{Qro.png}
    \caption{The graph of \(r \mapsto \int_0^{2\pi} a(r,\theta) d\theta\).}
    \label{fig:acrobot-Q}
\end{figure}
%
Notice that \(L\) is a positive constant which depends only on \(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(r,\theta)\) is adimensional. We have
\[
\int_0^{2\pi} \partial_I g(r,\theta,I) \big|_{I=0} d \theta =L \int_0^{2\pi} a(r,\theta) d\theta.
\]
The graph of \(\int_0^{2\pi} a(r,\theta) d\theta\), shown in
Figure \ref{fig:acrobot-Q}, indicates that for each $r \in \mathopen] 0,\pi\mathclose[$, the function $r \mapsto \int_0^{2\pi} a(r,\theta) d\theta$ is positive, and assumption~\eqref{eq:lemma:assumptions:3} holds. Thus the hypotheses of Lemma~\ref{lem:gain_lose} are satisfied, proving the theorem.
\hfill\QED

@@

In $(r,\theta)$ coordinates, both $T^+$ and $T^-$ map
system~\eqref{eqn:acrobot-constrained-dynamics} to a system of the
form~\eqref{eq:r_theta_system}, where one can verify that
\(f_r(r,\theta,0) \equiv 0\) and
\[
 f_\theta(r,\theta,0) = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
,
\]
which is positive so assumption~\eqref{eq:lemma_assumptions:1} holds. Letting $g
= f_r / f _\theta$, one can verify that \(\partial_r g (r,\theta,0)
\equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. The function
\(b(r,\theta) = \partial_I g(r,\theta,0)\) is given in the statement
of Theorem~\ref{thm:acrobot-rotations}, and by assumption we have that
condition~\eqref{eq:lemma:assumptions:3} of Lemma~\ref{lem:gain_lose} holds.
Theorem~\ref{thm:acrobot-rotations} now follows from  Lemma~\ref{lem:gain_lose}
and the fact that $\cU^+ \cup \cU^- = E_{>\bar R}$.
\hfill \QED
